Shade leaves of *Eucalpytus* trees lie in wait to maximize whole canopy carbon gain

### 2016-05-11 gm_newphyt/manuscript_gmes.Rmd
added clean aci files and restarted ACc script for manuscript
waiting for gmes curve fits
thesis reviewer comments on manuscript and thesis versions

---
title: Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

@@

\
```{r wc, echo=FALSE, warning=FALSE}
library(pixiedust)
wc_tab <- data.frame(count = c("Total word count (main body):", "Summary:", "Introduction:", "Materials and Methods:", "Results:",                               "Discussion:", "Acknowledgements:"), 
                     answer = c(5535, 219, 1207, 1847, 1102, 1359,20),
                     count2 = c("No. of Figures:", "No of Tables:", "No. of Supporting Information files:"
                                ,"","","",""),
                     answer2 = c("6 (all in color)",2,"6 (5 Figures, 1 Methods)", "","","",""))
#colnames(wc_tab) <- NULL 
dust(wc_tab)%>%
sprinkle_print_method("markdown")
```

@@

```{r global data/sourcing , include=FALSE}
source("functions and packages/functions.R")
source("functions and packages/packages.R")
source("master_scripts/plot_objects.R")
library(pixiedust)

@@

airvarfig <- "S1"
deltafig <- "S2"
accfig <- "S3"
get1fig <- "S4"
get2fig <- "S5"

@@

\
Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand]. This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~). Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Mesophyll conductance can also impose limitations on *A~n~* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast. Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified. Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.

@@

\
Once CO~2~ and water vapour flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400XT. The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL. Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and samples line gas streams for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle. Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.

\
Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:

@@

\
Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA). The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$. Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements. Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed. All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:

\
(3)
$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$

Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near]. Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].

@@

**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPa and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

@@

dust(resourcetab2)%>%
  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0("$\\delta$","^13^c")) %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```  
\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c). 
\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3**. The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals for the mean.
\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/physiology_boxplots.R")
```  
\
**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
source("master_scripts/wateruse.R")
```  
\
**Figure 5**. (a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
\
```{r gmshade, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gmes_shadeleaves_figure.R")
```  
\
**Figure 6**. The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree. Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased. The dashed line is the 1:1 relationship.

\
```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/cicc_boxplots.R")
```  
\
**Figure 7**. Boxplots of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).

### Acknowledgements
We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assistance.

@@

```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/air_vars.R")
```
\
**Figure S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
\
```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/delta_cica.R")
```  
\
**Figure S2**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light. The solid line represents the theoretical line for C3 plants from Evans et al. [-@evans1986carbon].
\
```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/ACc_figure.R")
```  
\
**Figure S3**. Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments. C~c~ values were predicted with g~m~ and curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).  
\
```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp3.R")
```  
\
**Figure S4**. Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively). 
\
```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp4.R")
```  
\
**Figure S5**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).

>>>>

---
title: Shade leaves of *Eucalpytus* trees lie in wait to maximize whole canopy carbon gain
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

@@

\
```{r wc, echo=FALSE, warning=FALSE}
library(pixiedust)
wc_tab <- data.frame(count = c("Total word count (main body):", "Summary:", "Introduction:", "Materials and Methods:", "Results:",                               "Discussion:", "Acknowledgements:"), 
                     answer = c(5535, 219, 1207, 1847, 1102, 1359,20),
                     count2 = c("No. of Figures:", "No of Tables:", "No. of Supporting Information files:"
                                ,"","","",""),
                     answer2 = c("6 (all in color)",2,"6 (5 Figures, 1 Methods)", "","","",""))
#colnames(wc_tab) <- NULL 
dust(wc_tab)%>%
sprinkle_print_method("markdown")
#Dynamic respones of shade leaf physiology to light availability increases whole tree carbon gain
```

@@

```{r install packages, echo=FALSE, message=FALSE, warning=FALSE,results="hide"}
#install needed packages if not already installed
.packages = c("lubridate", "doBy", "sciplot", "plyr","plotrix", "wesanderson", "plantecophys", "scales","reshape", "pixiedust")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], repos = "http://cran.at.r-project.org")

# Load packages into session
lapply(.packages, library, character.only=TRUE)
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/functions.R")
source("master_scripts/plot_objects.R")

@@

airvarfig <- "S1"
deltafig <- "S2"
get1fig <- "S3"
get2fig <- "S4"

@@

\
Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand]. This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~). Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Mesophyll conductance can also impose limitations on *A~n~* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast. Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified. It is likely that leaf anatomical costs associated with minimizing the length, diameter and tortuosity of the g~m~ diffusion pathway are necessary to maintain a high g~m~ [@hassiotou2009influence]. Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.

@@

\
Once CO~2~ and water vapour flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400XT. The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL. Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and samples line gas streams for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle. Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval. 
\
Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:

@@

\
Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA). The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$. Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements. Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed. All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:

\
(3)
$$K_l = \frac{E}{\Psi_{pd}~-\Psi_l}$$

Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E. The g~1~ parameter, describing the plants water-use strategy, was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near]. Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].

@@

**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

@@

dust(resourcetab2)%>%
  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K",paste0("$\\Psi$","~pre~"),paste0("$\\Psi$","~l~"), paste0("$\\delta$","^13^C")) %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/aci_vcmaxnitro.R")
```

\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3**. The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals for the mean.
\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/physiology_boxplots.R")
```  
\
**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
source("master_scripts/wateruse.R")
```  
\
**Figure 5**. (a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
\
```{r gmshade, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gmes_shadeleaves_figure.R")
```  
\
**Figure 6**. The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree. Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased. The dashed line is the 1:1 relationship.
\
```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/cicc_boxplots.R")
```  
\
**Figure 7**. Boxplots of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).

### Acknowledgements
We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assistance. We thank Sune Linder and the Swedish University for Agricultural Science for providing the whole-tree chambers.

@@

```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/air_vars.R")
```
\
**Figure S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
\
```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/delta_cica.R")
```  
\
**Figure S2**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light. The solid line represents the theoretical line for C3 plants from Evans et al. [-@evans1986carbon].
\
```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp3.R")
```  
\
**Figure S3**. Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively). 
\
```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp4.R")
```  
\
**Figure S4**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
